# Sample Size

Import the workspace established at the top levels script

```{r}
load(file = "C:/Users/DELL/Documents/socog/index.RData")
library(pracma)
library(ggplot2)
```

Estimate parameters for these simulations

```{r}
df = n1000
df = as.data.frame(df)
df$NLL = 0
df$NLL_alt = 0
for (i in 1:nrow(df)){
  row = which(round(freeParameters$theta, 2) == round(df$theta[i], 2) & 
                round(freeParameters$phi, 2) == round(df$phi[i], 2))
  preds = as.numeric(predictions[row, ])
  preds = rep(preds, times = nTrials/length(preds))
  preds = noise(preds)
  data = rbind(trialSet, trialSet)
  data$Returned = preds
  
  result = fmincon(obj_function,x0 = initial_params,
                   lb = lower_bounds, ub = upper_bounds,
                   df = data)
  resultAlt = fmincon(obj_functionAlt, x0 = initial_paramsAlt,
                      lb = lower_boundsAlt, ub = upper_boundsAlt,
                      df = data)
  data$Preds = generatePredictions(result$par, data)
  data$PredsAlt = generatePredictionsAlt(resultAlt$par, data)
  
  df$NLL[i] = -sum(dnorm(data$Returned, mean = data$Preds, 
                         sd = sd(data$Returned - data$Preds), log = TRUE))
  df$NLL_alt[i] = -sum(dnorm(data$Returned, mean = data$PredsAlt, 
                             sd = sd(data$Returned - data$PredsAlt), log = TRUE))
}
```

Compute BIC

```{r}
df$BIC = 2 * df$NLL + 2 * log(nrow(data))
df$BIC_alt = 2 * df$NLL_alt + 3 * log(nrow(data))
```

Now paired t-test over all values

```{r}
t.test(df$BIC, df$BIC_alt, paired = T)
```

Now we assess model goodness for each sample size

## ***N*** **= 20**

```{r}
sampleSize = 20
test = vector('logical', 100000)
better = vector('logical', 100000)
for (i in 1:length(test)){
  subsample = df[sample(1:nrow(df), sampleSize, replace = T), ]
  better[i] = sum(subsample$BIC) < sum(subsample$BIC_alt)
  test[i] = t.test(subsample$BIC, subsample$BIC_alt, paired = T)$p.value < 0.05
}
```

In 100,000 samples of 20 participants completing 60 trials with 85% of trials within 1 euro of the model's prediction, how often will we fail to identify the true model using BIC?

```{r}
summary(better)
```

And how many times will we show significant differences at a 95% CI?

```{r}
summary(test)
```

And both?

```{r}
contingency = table(better, test); contingency
```

```{r}
plottingInfoAll = data.frame(Outcome = rep(c("Incorrect & N.S.", 
                                             "Correct & N.S.", 
                                             "Incorrect & p < 0.05",
                                             "Correct & p < 0.05"), times = 5),
                             Count = 0,
                             N = rep(c(20, 50, 100, 200, 500), each = 4))
plottingInfoAll$Count[which(plottingInfoAll$N == sampleSize)] = 
  as.vector(contingency)
```

## ***N*** **= 50**

```{r}
sampleSize = 50
test = vector('logical', 100000)
better = vector('logical', 100000)
for (i in 1:length(test)){
  subsample = df[sample(1:nrow(df), sampleSize, replace = T), ]
  better[i] = sum(subsample$BIC) < sum(subsample$BIC_alt)
  test[i] = t.test(subsample$BIC, subsample$BIC_alt, paired = T)$p.value < 0.05
}
contingency = table(better, test)
plottingInfoAll$Count[which(plottingInfoAll$N == sampleSize)] = 
  as.vector(contingency)
```

## ***N*** **= 100**

```{r}
sampleSize = 100
test = vector('logical', 100000)
better = vector('logical', 100000)
for (i in 1:length(test)){
  subsample = df[sample(1:nrow(df), sampleSize, replace = T), ]
  better[i] = sum(subsample$BIC) < sum(subsample$BIC_alt)
  test[i] = t.test(subsample$BIC, subsample$BIC_alt, paired = T)$p.value < 0.05
}
contingency = table(better, test)
plottingInfoAll$Count[which(plottingInfoAll$N == sampleSize)] = 
  as.vector(contingency)
```

## ***N*** **= 200**

```{r}
sampleSize = 200
test = vector('logical', 100000)
better = vector('logical', 100000)
for (i in 1:length(test)){
  subsample = df[sample(1:nrow(df), sampleSize, replace = T), ]
  better[i] = sum(subsample$BIC) < sum(subsample$BIC_alt)
  test[i] = t.test(subsample$BIC, subsample$BIC_alt, paired = T)$p.value < 0.05
}
contingency = table(better, test)
plottingInfoAll$Count[which(plottingInfoAll$N == sampleSize)] = 
  as.vector(contingency)
```

## ***N*** **= 500**

```{r}
sampleSize = 500
test = vector('logical', 100000)
better = vector('logical', 100000)
for (i in 1:length(test)){
  subsample = df[sample(1:nrow(df), sampleSize, replace = T), ]
  better[i] = sum(subsample$BIC) < sum(subsample$BIC_alt)
  test[i] = t.test(subsample$BIC, subsample$BIC_alt, paired = T)$p.value < 0.05
}
contingency = table(better, test)
plottingInfoAll$Count[which(plottingInfoAll$N == sampleSize)] = 
  as.vector(contingency)
```

Now visualize all results

```{r}
plottingInfoAll$N = as.factor(plottingInfoAll$N)
plottingInfoAll$Outcome = factor(plottingInfoAll$Outcome,
                                 levels = c("Correct & p < 0.05",
                                            "Correct & N.S.", 
                                            "Incorrect & N.S.",
                                            "Incorrect & p < 0.05"))
plottingInfoAll$Proportion = plottingInfoAll$Count/100000
plottingInfoAll$Position = plottingInfoAll$Proportion/2

these = which(plottingInfoAll$Outcome == "Correct & p < 0.05")
plottingInfoAll$Position[these] = 1 - plottingInfoAll$Proportion[these]/2

those = which(plottingInfoAll$Outcome == "Correct & N.S.")
plottingInfoAll$Position[those] = 1 - plottingInfoAll$Proportion[these] -
  plottingInfoAll$Proportion[those]/2

these = which(plottingInfoAll$Outcome == "Incorrect & p < 0.05")
those = which(plottingInfoAll$Outcome == "Incorrect & N.S.")
plottingInfoAll$Position[those] = plottingInfoAll$Proportion[these] +
  plottingInfoAll$Proportion[those]/2
ggplot(data = plottingInfoAll, aes(x = Proportion, y = N, group = Outcome,
                                   fill = Outcome)) + 
  geom_col() + theme_minimal() + 
  scale_fill_manual(values = c("green", "forestgreen", "firebrick4", "red")) +
  labs(x = "Proportion", y = "Sample Size", 
       fill = "Result",
       label = NULL) + 
  geom_label(aes(label = Proportion, x = Position)) + lims(x = c(-0.05, 1.05))
```

And save for other uses and verification

```{r}
save.image(file = "C:/Users/DELL/Documents/socog/sampleSize.RData")
```
